Comparison of Fast Imaging Architectures for Large, Multi-scale, Hierarchical Aperture Arrays

### 2024-06-02 ms.tex
Update on Overleaf.
Update correlator beamformer section

The computational cost budget consists of the following components. The formation of polarimetric visibilities and their accumulations requires $n_\textrm{p}^2 N_\textrm{a/s} (N_\textrm{a/s}-1)/2$ complex multiplications and complex additions, respectively, per station per channel at a cadence of $\delta t$. These accumulated visibilities will be further averaged across stations but on a slower cadence of $t_\textrm{acc}$. However, in order to obtain stationwise image intensities, we will not consider averaging across stations. Thus, the computational cost of forming polarimetric visibilities and their accumulations per station per channel is $4 n_\textrm{p}^2 N_\textrm{a/s} (N_\textrm{a/s}-1)/(\delta t)$ FLOPS. The accumulated visibilities from a station can be beamformed towards $n_\textrm{bs}\simeq (D_\textrm{s}/D_\textrm{a})^2$ sky locations using a DFT to obtain Stokes intensities. This DFT beamforming performs $n_\textrm{p}^2 N_\textrm{a/s} (N_\textrm{a/s}-1) (D_\textrm{s}/D_\textrm{a})^2 /2$ CMACs at a slower cadence of $t_\textrm{acc}$. Hence, the net computational cost for DFT beamforming of visibilities across the field of view per station per channel is 
\begin{align}
    C_\textrm{XBF} &= 4 n_\textrm{p}^2 \, N_\textrm{a/s} (N_\textrm{a/s}-1) \left[\frac{1}{\delta t} + \frac{\left(D_\textrm{s}/D_\textrm{a}\right)^2}{t_\textrm{acc}} \right] \, . \label{eqn:cost-XBF}
\end{align}

\begin{figure}
\includegraphics[width=\linewidth]
% {figures/1D_xcor_dft_cost_analysis_SKA1-low.pdf}
{figures/1D_xcor_dft_per_pixel_cost_analysis_LAMBDA-I.pdf}
% \includegraphics[width=\linewidth]{fig_01.pdf}
\caption{
\label{fig:1D-incoherent-compcost-XBF-LAMBDA-I}}
\end{figure}

\begin{figure*}
\centering
\subfloat[][Two-dimensional slices \label{fig:multidim-incoherent-compcost-XBF-LAMBDA-I}]
{\includegraphics[width=0.48\textwidth]
% {figures/multidim_xcor_dft_cost_analysis_SKA1-low.pdf}
{figures/multidim_xcor_dft_per_pixel_cost_analysis_LAMBDA-I.pdf}
}
\subfloat[][Two-dimensional slices (relative)\label{fig:multidim-incoherent-relative-compcost-XBF-LAMBDA-I}]
{\includegraphics[width=0.48\textwidth]
% {figures/multidim_xcor_dft_relative_cost_analysis_SKA1-low.pdf}
{figures/multidim_xcor_dft_relative_cost_analysis_LAMBDA-I.pdf}
}
\caption{(Left): Two-dimensional slices of the computational cost for coherent-incoherent imaging using a DFT of station-level cross-correlations for \texttt{SKA-low-core}, \texttt{SKA-low}, \texttt{Aus-VLBA-low}, and \texttt{IC-VLBA-low}. Cyan contours  \label{fig:incoherent-compcost-XBF-LAMBDA-I}}
\end{figure*}

>>>>

Figure~\ref{fig:1D-incoherent-compcost-XBF-LAMBDA-I} shows the computational cost density breakdown for correlator beamforming at the station level. The correlation and temporal averaging in Equation~(\ref{eqn:intra-station-pol-visibilities}) scale as $N_\textrm{a/s}(N_\textrm{a/s}-1)/2$ and has no dependence on $D_\textrm{a}$ or $D_\textrm{s}$. Hence, their per pixel cost scales as $(D_\textrm{s}/D_\textrm{a})^{-2}$. The two-dimensional DFT in Equation~(\ref{eqn:intra-station-pol-xbf-img-expl}) also scales as $N_\textrm{a/s}(N_\textrm{a/s}-1)/2$, but being a per-pixel operation, stays constant with $D_\textrm{a}$ and $D_\textrm{s}$. The correlation and temporal averaging have to be performed at every interval of $\delta t$ and are therefore, independent of $t_\textrm{acc}$, whereas the DFT needs to be performed at a cadence of $t_\textrm{acc}$ and thus scales as $t_\textrm{acc}^{-1}$. For the chosen parameters of \texttt{LAMBDA-I} and $t_\textrm{acc}=1$~ms, the computational cost density is dominated by the cost of the DFT. However, for imaging on a slower cadence, $t_\textrm{acc}\gtrsim 10$~ms, the cost of correlation begins to dominate.

\begin{figure}
\includegraphics[width=\linewidth]
% {figures/1D_xcor_dft_cost_analysis_SKA1-low.pdf}
{figures/1D_xcor_dft_per_pixel_cost_analysis_LAMBDA-I.pdf}
% \includegraphics[width=\linewidth]{fig_01.pdf}
\caption{A breakdown of the computational cost density function over the station parameters for the correlator beamforming architecture (XBF) at the station level. Each panel shows the variation with the respective parameter keeping the rest fixed at the characteristic values of the \texttt{LAMBDA-I}, \texttt{SKA-low-core}, and \texttt{SKA-low} stations (cyan lines). The two-dimensional DFT (dashed lines) dominates the cost for the chosen \texttt{LAMBDA-I} parameters at an imaging cadence of $t_\textrm{acc}=1$~ms over correlation (double dot-dashed) and temporal averaging of the correlations (dot-dashed).
% The spatial FFT cost in Equation~(\ref{eqn:intra-station-pol-hol-img-epic-LA}) denoted by black dashed lines dominates the squaring (pink dot-dashes) and temporal averaging (blue dot-dashes) costs in Equation~(\ref{eqn:intra-station-opt-polarimetric-direct-img-LA}). The gridding operation (black dotted lines) in Equation~(\ref{eqn:intra-station-pol-hol-img-epic-LA}) contributes even lesser to the cost budget. 
\label{fig:1D-incoherent-compcost-XBF-LAMBDA-I}}
\end{figure}

Figure~\ref{fig:multidim-incoherent-compcost-XBF-LAMBDA-I} shows the covariance of the total computational cost density per station per voxel for the correlator beamformer architecture taken two station parameters at a time while keeping the rest fixed at the nominal values of a \texttt{LAMBDA-I} station denoted by grey dashed lines. The variation in the cost per voxel for $t_\textrm{acc}=1$~ms is dominated by the spatial DFT cost which is relatively constant in $D_\textrm{a}$ and $D_\textrm{a}$ and scales as $N_\textrm{a/s}^2$, displaying the same trends observed in Figure~\ref{fig:1D-incoherent-compcost-XBF-LAMBDA-I}. Figure~\ref{fig:multidim-incoherent-relative-compcost-XBF-LAMBDA-I} shows the ratio of the computational cost of the XBF architecture relative to the voltage beamforming architecture in Figure~\ref{fig:multidim-incoherent-compcost-BF-LAMBDA-I}. It is clearly noted that the cost of the XBF architecture is less expensive only for $N_\textrm{a/s}\lesssim 100$ and gets worse with increasing $N_\textrm{a/s}$.

\begin{figure*}
\centering
\subfloat[][Two-dimensional slices \label{fig:multidim-incoherent-compcost-XBF-LAMBDA-I}]
{\includegraphics[width=0.48\textwidth]
% {figures/multidim_xcor_dft_cost_analysis_SKA1-low.pdf}
{figures/multidim_xcor_dft_per_pixel_cost_analysis_LAMBDA-I.pdf}
}
\subfloat[][Two-dimensional slices (relative)\label{fig:multidim-incoherent-relative-compcost-XBF-LAMBDA-I}]
{\includegraphics[width=0.48\textwidth]
% {figures/multidim_xcor_dft_relative_cost_analysis_SKA1-low.pdf}
{figures/multidim_xcor_dft_relative_cost_analysis_LAMBDA-I.pdf}
}
\caption{(Left): Two-dimensional slices of the computational cost for coherent-incoherent imaging using a DFT of station-level cross-correlations for \texttt{LAMBDA-I}, \texttt{SKA-low-core}, and \texttt{SKA-low} station parameters (dashed gray lines). Cyan lines denote contours of the colour scale in logarithmic increments. (Right): Same as the left but relative to a voltage beamformer architecture. The computational cost density is dominated by the two-dimensional DFT cost for $t_\textrm{acc}=1$~ms, and is lower than that of voltage beamforming for $N_\textrm{a/s}\lesssim 100$ while getting more expensive with increasing $N_\textrm{a/s}$. \label{fig:incoherent-compcost-XBF-LAMBDA-I}}
\end{figure*}
